\bf The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-05-13 paper/paper.Rmd
Dave future work edit

The `forestecology` package provides an accessible way to fit and test models of neighborhood competition. While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken. Furthermore, we also envision the package being used more generally for inventory plot-structure with many spatially separated plots, e.g. the US Forest Service Forest Inventory and Analysis plots [@smith_forest_2002]. In future versions of `forestecology` we also hope to include models that account for tree mortality in addition to tree growth. The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure. We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.

>>>>

The `forestecology` package provides an accessible way to fit and test models of neighborhood competition. The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure. We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.

While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken. Furthermore, we hope that future versions of the package will be flexible to other plot layouts, for example inventory plot-structure with many spatially separated plots like the US Forest Service Forest Inventory and Analysis plots [@smith_forest_2002].

We also hope to extend the `forestecology` package's functionality to account for a larger variety of models for tree growth. One clear future direction would be to allow competition based on species trait values rather than species identity. There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012, @lasky_trait-mediated_2014, @uriarte_trait_2010]. Thus an extenstion of the model would allow $lambda$ values to be a function of the traits of competing species.

Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \ref{spatial-information} and \ref{spatial-cross-validation} [@valavi_blockcv_2019]. This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package, @tidymodels_package].

### 2021-05-13 paper/paper.tex
Dave future work edit

\caption{Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates. The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations. The colors indicate whether cross validaton was used.}\label{fig:scbi-simulation}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The \texttt{forestecology} package provides an accessible way to fit and
test models of neighborhood competition. While the package is designed
with ForestGEO plot data in mind, we envision that it can be modified to
work on any single large, mapped forest plot in which at least two
measurements of each individual have been taken. Furthermore, we also
envision the package being used more generally for inventory
plot-structure with many spatially separated plots, e.g.~the US Forest
Service Forest Inventory and Analysis plots \citep{smith_forest_2002}.
In future versions of \texttt{forestecology} we also hope to include
models that account for tree mortality in addition to tree growth. The
package follows the \texttt{tidy} data design principles, leverages the
\texttt{sf} package for spatial data, and S3 open-oriented model
implementation structure. We hope that the package will increase the use
of neighborhood competition models to better understand what structures
plant competition.

>>>>

\caption{Comparison of root mean squared error of models for standard, permuted, and spatially cross-validated error estimates. The dotted lines show observed RMSE while the histograms show the null distribution of RMSE for 49 permutations under the null hypothesis of no competitor species identity effects. The colors indicate whether spatial cross-validation was used or not.}\label{fig:scbi-simulation}
\end{figure}

\hypertarget{conclusion-and-future-work}{%
\section{Conclusion and future work}\label{conclusion-and-future-work}}

The \texttt{forestecology} package provides an accessible way to fit and
test models of neighborhood competition. The package follows the
\texttt{tidy} data design principles, leverages the \texttt{sf} package
for spatial data, and S3 open-oriented model implementation structure.
We hope that the package will increase the use of neighborhood
competition models to better understand what structures plant
competition.

While the package is designed with ForestGEO plot data in mind, we
envision that it can be modified to work on any single large, mapped
forest plot in which at least two measurements of each individual have
been taken. Furthermore, we hope that future versions of the package
will be flexible to other plot layouts, for example inventory
plot-structure with many spatially separated plots like the US Forest
Service Forest Inventory and Analysis plots \citep{smith_forest_2002}.

We also hope to extend the \texttt{forestecology} package's
functionality to account for a larger variety of models for tree growth.
One clear future direction would be to allow competition based on
species trait values rather than species identity. There is evidence
that traits predict competitive outcomes
\citep[\citet{lasky_trait-mediated_2014},
\citet{uriarte_trait_2010}]{kunstler_competitive_2012}. Thus an
extenstion of the model would allow \(lambda\) values to be a function
of the traits of competing species.

Lastly, the \texttt{forestecology} current uses the \texttt{blockCV}
package behind the scenes to create the spatial blocks acting as folds
for our spatial cross-validation algorithm detailed in Sections
\ref{spatial-information} and \ref{spatial-cross-validation}
\citep{valavi_blockcv_2019}. This back-end functionality could be
substituted with the \texttt{spatialsample} package for spatial
resampling infrastructure; a \texttt{tidymodels} package under active
development as of 2021
\citep[\citet{tidymodels_package}]{spatialsample_package}.
